Hierarchical Principal Manifold Estimation

### 2023-10-03 paper/main.tex
edits

We begin this section with a brief review of additive models. Consider a regression scenario where the dependent variable $\mathbf{y} = \left(y_1, \dots, y_n\right)^{T}$ is thought to be associated with independent covariate vectors $\mathbf{x}_i = \left(x_{i1}, \dots, x_{ip}\right)$. In this situation, the standard multiple linear regression model assumes that the dependence of $y_i$ on $\mathbf{x}_i$ takes the linear form
\[%
  y_i = \alpha + x_{i1}\beta_1 + \dots + x_{ip}\beta_p + \epsilon_i
,\]%
where $\epsilon$ has mean 0 and variance $\sigma^2$. Additive models are a direct extension of this model, where the additive structure of the model is retained, but the assumption of linearity is discarded. A simple additive model takes the form
\begin{equation}
  y_i = \alpha + \sum_{j=1}^{p}f_j(X_{ij}) + \epsilon_i, \label{eq:1}
\end{equation}
where again $\epsilon_i$ has mean 0 and variance $\sigma^2$, however for each $j$, $f_j$ represents an arbitrary function commonly assumed to be smooth. While the functions in model \ref{eq:1} only take as parameters the value of one covariate, with each covariate being represented in the model once, this does not need to be the case. The structure of additive models seeks to balance benefits of interpretability found in additive models with the flexibility of nonlinear, and frequently, nonparametric functional forms. Further information can be found in \cite{hastieGeneralizedAdditiveModels1990}.

@@

Individual participants are typically scheduled for imaging at six month intervals, though variability exists in terms of the recorded intervals. Also of note, the numbers of observations contributed by each participant varies, so groups may be unbalanced. We denote the number of groups under consideration using $G$, the number of individuals included in the study as $I$, the number of images available as $N$, and the number of observations available for the $j$th image, with $j = 1, 2, \dots, N$, as $M_j$. The individual to which the $j$th image belongs is denoted as $i(j)$, the group of the $i(j)$th individual is represented by $g(j)$, and we use $t_j$ to denote the time at which the $j$th image was recorded, with all times being relative to the $i(j)$th individual's baseline visit. Finally, we use $N_t$, $N_i$, and $N_g$ to denote the number of images available for time point $t$, individual $i$, and group $g$, respectively.

Within a segmented structural MRI, the voxels estimated to mark the surface of the hippocampus can be assumed to lie along a two-dimensional manifold that has been wrapped into the three-dimensional image space. More generally, we let $d$ denote the intrinsic dimension of the low-dimensional manifold, and $D$ represent the dimension of the high-dimensional input space. For the $j$th image with $j = 1, \dots, N$, we denote the $m$th voxel in the $D$-dimensional space as $\mathbf{x}_{jm} \in \mathbb{R}^{D}$, its associated $d$-dimensional parameterization as $\mathbf{y}_{jm} \in \mathbb{R}^{d}$, and its weight as $w_{jm}$. We use $\mathbf{x}_g$, $\mathbf{x}_i$, $\mathbf{x}_{j}$, and $\mathbf{x}_t$ to denote the input voxels from all images in the $g$th group, $i$th individual, $j$th image, and $t$th time point, respectively, and use similar notation for all other relevant terms.

The embedding function which maps from the $d$-dimensional manifold to $D$-dimensional ambient space is represented as $\psi(\mathbf{y}_{jm})$, which can be modeled as 
\begin{equation}
  \mathbf{x}_{jm} = \psi(\mathbf{y}_{jm}, t_j) = f_p(\mathbf{y}_{jm}, t_{j}) + f_{g(j)}(\mathbf{y}_{jm}, t_{j}) + f_{i(j)}(\mathbf{y}_{jm}, t_{j}) + f_j(\mathbf{y}_{jm}) + \epsilon_{jm}. \label{eq:2}
\end{equation}
The embedding map $\psi(\mathbf{y})$ is associated with a projection index $\kappa_{\psi}(\mathbf{x}): \mathbb{R}^{D} \to \mathbb{R}^{d}$, which, informally, outputs the $d$-dimensional parameterization $\mathbf{y}$ that minimizes $\|\mathbf{x} - \psi(\mathbf{y})\|_{\mathbb{R}^{D}}$ and follows the generalized definition of projection indices given in \cite{mengPrincipalManifoldEstimation2021}.

@@

The smoothing spline functions $f(\mathbf{y})$ minimize 
\[%
  \mathcal{K}_{\lambda}(f) = \sum_{j=1}^{N}\sum_{m=1}^{M_j}w_{jm}\|\mathbf{x}_{jm} - f(\mathbf{y}_{jm})\|^2 + \lambda\|\nabla^{\otimes 2}f\|^2
,\]%
where $\nabla^{\otimes 2}f$ denotes the Hessian matrix of $f$, and $\lambda$ is given. Thus, $f(\mathbf{y})$ takes the form
\begin{equation}
  f_l(\mathbf{y}) = \sum_{j=1}^{N}\sum_{m=1}^{M_j}s_{jm, l} \times \eta_{d}\left(\|\mathbf{y} - \mathbf{y}_{jm}\|\right) + \sum_{p = 1}^{d + 1}\alpha_{p, l} \times \phi_{p}(\mathbf{y}), \quad l = 1, 2, \dots, D, \label{eq:3}
\end{equation}
with $\phi_p(\mathbf{y})$ representing the basis of the linear space of polynomials on $\mathbb{R}^{d}$ with degree $\leq 1$ and $\eta_{d}(y) = y^{4-d}\log y$ when $\log y \neq 0$ and $d$ is even, $\eta_{d}(y) = 0$ when $\log y = 0$ and $d$ is even, and $\eta_{d}(y) = y^{4-d}$ when $d$ is odd.

Without accounting for the additive structure of the model, and after concatenating all relevant values into $M = \sum_{j=1}^{N}M_{j}$-dimensional vectors of input values $\mathbf{y}$, weights $\mathbf{w}$, and outputs $\mathbf{x}$, we can estimate $f$ by obtaining the values of $s_l \in \mathbb{R}^{M}$ and $\alpha_l \in \mathbb{R}^{d + 1}$ for $l = 1, \dots, D$ that minimize $\|\mathbf{W}^{\frac{1}{2}}\left(\mathbf{x}_{l} - \mathbf{E}s_l - \mathbf{R}\alpha_l\right)\|^2_{\mathbb{R}^{M}} + \lambda\|\mathbf{E}^{\frac{1}{2}}s_l\|^2_{\mathbb{R}^{M}}$ under the constraint that $\mathbf{R}^{T}s_l = 0$, where
\begin{itemize}
  \item $\mathbf{E} = (E_{qr})_{1 \leq q, r, \leq M}$ is defined by $E_{qr} = \eta_{d}\left(\|\mathbf{y}_{q} - \mathbf{y}_{r}\|\right))$.
  \item $\mathbf{R} = \left(R_{mp}\right)_{1 \leq m \leq M, \ 1 \leq p \leq d + 1}$ is defined by $R_{mp} = \phi_p(\mathbf{y}_{m})$.
  \item $\mathbf{W} = diag(w_{1}, \dots, w_{M})$.
\end{itemize}

@@

\begin{equation}
   \left(\begin{array}{ccc}
    2\mathbf{EWE} + 2\lambda\mathbf{E} & 2\mathbf{EWR} & \mathbf{R} \\
    2\mathbf{R}^{T}\mathbf{WE} & 2\mathbf{R}^{T}\mathbf{WR} & \mathbf{0} \\
    \mathbf{R}^{T} & \mathbf{0} & \mathbf{0}
  \end{array}\right)\left(
  \begin{array}{c}
    s_l \\
    \alpha_l \\
    \upsilon_l
  \end{array}
  \right) = \left(
  \begin{array}{c}
    2\mathbf{EW}\mathbf{x}_l \\
    2\mathbf{R}^{T}\mathbf{W}\mathbf{x}_l \\
    \mathbf{0}
  \end{array}
  \right). \label{eq:4}
\end{equation}
The optimal value of $\lambda$ can be found using cross-validation, where with images being randomly split into $K$ folds, functions may be evaluated by the mean squared error, written as
\begin{equation}
  MSE(\lambda) = \frac{1}{K}\sum_{k=1}^{K}\frac{1}{N_u}\sum_{j=1}^{N_u}\frac{1}{M_j}\sum_{m=1}^{M_j}\|\mathbf{x}_{jm} - f_{\lambda}^{(k)}(\mathbf{y}_{jm})\|^2. \label{eq:5}
\end{equation}
Here, $f_{\lambda}^{(k)}(\mathbf{y})$ denotes the estimated spline function given tuning parameter $\lambda$ and with the data for the $k$th fold excluded.

The coefficients and tuning parameter that minimize $\mathcal{K}_{\lambda}(f)$ suffice for estimating $f_{j}(\mathbf{y})$, in which the coefficients are not time-varying. To reach estimates for the varying-coefficient models, equation \ref{eq:4} is solved using the data available at each time point within the relevant study level. To estimate $f_p(\mathbf{y}, t)$, we use $\left\{\mathbf{x}_t\right\}_{t=1}^T$ and $\left\{\mathbf{y}_t\right\}_{t=1}^T$, yielding fitted function $\left\{\hat{f}_t(\mathbf{y})\right\}_{t=1}^T$. Using the maximum and minimum values of $\mathbf{y}_t$ in each dimension, a new $d$-dimensional grid of $M' = \text{max}\left(\left\{M_{j}\right\}_{j = 1}^{N}\right)$ knots, $\left\{\mathbf{y'}_m\right\}_{m=1}^{M'}$, is generated with the knots being evenly spread throughout the parameter space. These parameters serve as common knots that allow the spline coefficients to be compared across time points. They are then used to compute a new set of corresponding outcome values $\left\{\mathbf{x}'_t\right\}_{t=1}^{T}$. The spline functions are reestimated as $\left\{\hat{f}'_t\right\}_{t=1}^{T}$ using $\left\{\mathbf{x}'_t\right\}_{t=1}^{T}$ and $\mathbf{y}'$. These newly generated knots are weighted equally, so an unweighted smoothing spline model is used. Following \cite{greenSilverman1994}, we let

\begin{itemize}
  \item $\mathbf{E}' = \left(E'_{qr}\right)_{1 \leq q, r, \leq M'}$ defined by $E'_{qr} = \eta_{d}\left(\|\mathbf{y}'_{q} - \mathbf{y}'_{r}\|\right)$.
  \item $\mathbf{R}' = \left(R'_{mp}\right)_{1 \leq m \leq M', \ 1 \leq p \leq d + 1}$ is defined by $R'_{mp} = \phi_p(\mathbf{y'}_{m})$.
\end{itemize}

@@

As with the tuning process for $\lambda$, the optimal value of $\gamma$ can be identified using cross-validation, with either leave-one-out or $k$-folds cross-validation being appropriate depending on the number of time points available. If using leave-one-out cross-validation, the mean squared error can be calculated as
\begin{equation}
  MSE(\gamma) = \frac{1}{T}\sum_{t=1}^{T}\frac{1}{N_t}\sum_{j=1}^{N_t}\frac{1}{M_j}\sum_{m=1}^{M_j}\|\mathbf{x}_{jm} - f_{\gamma}^{(t)}(\mathbf{y}_{jm}, t_j)\|^2. \label{eq:8}
\end{equation}
We then let $\hat{\gamma}$ be the value of $\gamma$ that minimizes $MSE(\gamma)$.

With a proper estimate of the $\hat{\gamma}$, we may denote the coefficients of $g(t)$ as $\Omega = \left(\delta^{T}, \tau^{T}\right)$, and those estimated by $g(t)$ as $\omega(t) = \left(\mathbf{s}_{\hat{\gamma}}(t)^{T}, \mathbf{\alpha}_{\hat{\gamma}}(t)^{T}\right)^{T}$. We use these coefficients in equation \ref{eq:3} in place of the constant coefficients. Therefore, the varying-coefficient models have the following outputs:
\begin{equation}
  f(\mathbf{y}, t) = \sum_{m=1}^{M'}\mathbf{s}_{\hat{\gamma}, l}(t) \eta_{d}\left(\|\mathbf{y} - \mathbf{y'}_m\|\right) + \sum_{p=1}^{d + 1}\mathbf{\alpha}_{\hat{\gamma}, p, l}(t) \phi_p(\mathbf{y}), \quad l = 1, \dots, D. \label{eq:9}
\end{equation}
The fitting process for the varying-coefficient smoothing spline model is given formally as Algorithm \ref{alg:varying_spline}.

@@

\begin{algorithm}
\caption{Varying-Coefficient Smoothing Spline Model}\label{alg:varying_spline}
  \KwData{$D$-dimensional output values $\left\{\mathbf{x}_{jm}\right\}_{j = 1, m = 1}^{N, M_{j}}$, $d$-dimensional input values $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, weights $\left\{w_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, observation times $\left\{t_{j}\right\}_{j=1}^{N}$, candidate tuning parameters $\left\{\lambda_l\right\}_{l=1}^{L}$, $\left\{\gamma_p\right\}_{p=1}^{P}$, and number of cross-validation folds $K$.}
  \KwResult{Analytic formula of $\hat{f}(\mathbf{y}, t)$, coefficients $\Omega = \left(\mathbf{\delta}^{T}, \mathbf{\tau}^{T}\right)$, and optimal tuning parameter $\hat{\gamma}$.}

\For{$t = 1, 2, \dots, T$} {
  Set $U_t$ equal to the number of images included in the data at time $t$\;
  \For{$l = 1, 2, \dots, L$} {
    Solve equation \ref{eq:4} to estimate $f_{\lambda_l, t}(\mathbf{y})$ with inputs $\mathbf{x}_t$, $\mathbf{y}_t$, and $w_t$\;
    \For{$u = 1, 2, \dots, U_t$} {
      Solve equation \ref{eq:4} to estimate $f^{(-u)}_{\lambda_l, t}(\mathbf{y})$ with inputs $\mathbf{x}^{(-u)}_t$, $\mathbf{y}^{(-u)}_t$, $w^{(-u)}_t$\;
      Estimate $MSE(\lambda_{l})_u$ using equation \ref{eq:5}\;
    }
    Set $MSE(\lambda_l) = \frac{1}{U}\sum_{u=1}^{U}MSE(\lambda_l)_u$\;
  }
  $\hat{\lambda}_t = \arg \min_{\lambda} MSE(\lambda)$\;
  $f_t(\mathbf{y}) = f_{\hat{\lambda}_t}(\mathbf{y})$\;
}

  Let $M' = \max(M_{j})$, $\left\{\mathbf{y}'_{m}\right\}_{m=1}^{M'}$ be a grid spanning the range of values of $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N, M_j}$\;
  \For{$t = 1, 2, \dots, T$} {
    Compute $\left\{\mathbf{x'}_{m, t}\right\}_{m=1}^{M'} = \left\{f_t(\mathbf{y'}_m)\right\}_{m=1}^{M'}$\;
    Solve equation \ref{eq:6} to compute $\mathbf{s'}_t$ and $\mathbf{\alpha'}_t$\;
    Set $\mathbf{\omega}_t = \left[\mathbf{s'}_t^{T}, \mathbf{\alpha'}_t^{T}\right]$\;
  }
  \For{$p = 1, 2, \dots, P$} {
    Compute $g_{\gamma_p}(t)$ by solving equation \ref{eq:7}\;
    \For{$t = 1, 2, \dots, T$} {
      Compute $g_{\gamma_p}^{(t)}$ by solving \ref{eq:7}\;
    }
    Estimate $MSE(\gamma_p)$ using \ref{eq:8}\;
  }
  $\hat{\gamma} = \arg \min_{\gamma} MSE(\gamma)$\;
  $\hat{f}(\mathbf{y}, t) = f_{\hat{\gamma}}(\mathbf{y}, t)$, where the form of $\hat{f}(\mathbf{y}, t)$ is given in \ref{eq:9}.
\end{algorithm}

@@

\[%
  f_p(\mathbf{y}_{jm}, t_{j}) = E\left[\mathbf{x}_{jm} - f_{g(j)}(\mathbf{y}_{jm}, t_{j}) - f_{i(j)}(\mathbf{y}_{jm}, t_{j}) - f_{j}(\mathbf{y}_{jm})\right]
,\]%
with a similar result holding for each of the smooth functions making up the additive model. Thus, the backfitting algorithm to fit the proposed model is given in Algorithm \ref{alg:backfitting}.

\begin{algorithm}
\footnotesize
\caption{Backfitting Algorithm for Additive Spline Model}\label{alg:backfitting}
  \KwData{$D$-dimensional output values $\left\{\mathbf{x}_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, $d$-dimensional input values $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, observation times $\left\{t_{j}\right\}_{j=1}^{N}$, weights $\left\{w_{jm}\right\}_{j=1, m=1}^{N, M_j}$, group and individual indicator values $\left\{g(j)\right\}_{j=1}^{N}$ and $\left\{i(j)\right\}_{j=1}^{N}$, candidate tuning parameters $\left\{\lambda_l\right\}_{l=1}^{L}$ and $\left\{\gamma_p\right\}_{p=1}^{P}$, $\epsilon$ and $K$.}
  \KwResult{Analytic formulas of $\hat{f}_p$, $\left\{\hat{f}_g\right\}_{g=1}^{G}, \left\{\hat{f}_i\right\}_{i=1}^{I}$, and $\left\{\hat{f}_{j}\right\}_{j=1}^{N}$.}

>>>>

We begin this section with a brief review of additive models. Consider a regression scenario where the dependent variable $\mathbf{y} = \left(y_1, \dots, y_n\right)^{T}$ is thought to be associated with independent covariate vectors $\mathbf{x}_i = \left(x_{i1}, \dots, x_{ip}\right)$, for $i = 1, \dots, n$, with $n$ representing the number of observations. In this situation, the standard multiple linear regression model assumes that the dependence of $y_i$ on $\mathbf{x}_i$ takes the linear form
\[%
  y_i = \alpha + x_{i1}\beta_1 + \dots + x_{ip}\beta_p + \epsilon_i
,\]%
where $\epsilon$ has mean 0 and variance $\sigma^2$. Additive models are a direct extension of this model, where the additive structure of the model is retained, but the assumption of linearity is discarded. A simple additive model takes the form
\begin{equation}
  y_i = \alpha + \sum_{j=1}^{p}f_j(X_{ij}) + \epsilon_i, \label{eq:1}
\end{equation}
where again $\epsilon_i$ has mean 0 and variance $\sigma^2$, however for each $j$, $f_j$ represents an arbitrary function commonly assumed to be smooth. While the functions in model \ref{eq:1} only take as parameters the value of one covariate, with each covariate being represented in the model once, this does not need to be the case, as functions specifying interactions between covariates may also be included. The structure of additive models seeks to balance benefits of interpretability found in linear regression models with the flexibility of nonlinear, and frequently, nonparametric functional forms. Further information can be found in \cite{hastieGeneralizedAdditiveModels1990}.

@@

Individual participants are typically scheduled for imaging at six month intervals, though variability exists in the recorded intervals. Also of note, the numbers of observations contributed by each participant varies, so groups may be unbalanced. We denote the number of groups under consideration using $G$, the number of individuals included in the study as $I$, the number of images available as $N$, and the number of observations available for the $j$th image as $Q_j$, with $j = 1, 2, \dots, N$. The individual to which the $j$th image belongs is denoted as $i(j)$, the group of the $i(j)$th individual is represented by $g(j)$, and we use $t_j$ to denote the time at which the $j$th image was recorded, with all times being relative to the $i(j)$th individual's baseline visit. This time benchmark may be changed to age, time from diagnosis, time from symptom onset, or any other measure of time, depending on the setting. Finally, we use $N_t$, $N_i$, and $N_g$ to denote the number of images available for time point $t$, individual $i$, and group $g$, respectively.

Within a segmented structural MRI, the voxels estimated to mark the surface of the hippocampus can be assumed to lie along a two-dimensional manifold that has been wrapped into the three-dimensional image space. More generally, we let $d$ denote the intrinsic dimension of the low-dimensional manifold, and $D$ represent the dimension of the high-dimensional input space. For the $j$th image with $j = 1, \dots, N$, we denote the $q$th voxel in the $D$-dimensional space as $\mathbf{x}_{jq} \in \mathbb{R}^{D}$, its associated $d$-dimensional parameterization as $\mathbf{y}_{jq} \in \mathbb{R}^{d}$, and its weight as $w_{jq}$. We use $\mathbf{x}_g$, $\mathbf{x}_i$, $\mathbf{x}_{j}$, and $\mathbf{x}_t$ to denote the input voxels from all images in the $g$th group, $i$th individual, $j$th image, and $t$th time point, respectively, and use similar notation for all other relevant terms.

The embedding function which maps from the $d$-dimensional manifold to $D$-dimensional ambient space is represented as $\psi(\mathbf{y}_{jq})$, which can be modeled as 
\begin{equation}
  \mathbf{x}_{jq} = \psi(\mathbf{y}_{jq}, t_j) = f_p(\mathbf{y}_{jq}, t_{j}) + f_{g(j)}(\mathbf{y}_{jq}, t_{j}) + f_{i(j)}(\mathbf{y}_{jq}, t_{j}) + f_j(\mathbf{y}_{jq}) + \epsilon_{jq}. \label{eq:2}
\end{equation}
The embedding map $\psi(\mathbf{y})$ is associated with a projection index $\kappa_{\psi}(\mathbf{x}): \mathbb{R}^{D} \to \mathbb{R}^{d}$, which, informally, outputs the $d$-dimensional parameterization $\mathbf{y}$ that minimizes $\|\mathbf{x} - \psi(\mathbf{y})\|_{\mathbb{R}^{D}}$ and follows the generalized definition of projection indices given in \cite{mengPrincipalManifoldEstimation2021}.

@@

The smoothing spline functions $f(\mathbf{y})$ minimize 
\[%
  \mathcal{K}_{\lambda}(f) = \sum_{j=1}^{N}\sum_{q=1}^{Q_j}w_{jq}\|\mathbf{x}_{jq} - f(\mathbf{y}_{jq})\|^2_{\mathbb{R}^{d}} + \lambda\|\nabla^{\otimes 2}f\|^2_{L^2(\mathbb{R}^{d})}
,\]%
where $\nabla^{\otimes 2}f$ denotes the Hessian matrix of $f$, and $\lambda$ is a given parameter used to penalize curvature in $f$. Thus, $f(\mathbf{y})$ takes the form
\begin{equation}
  f_l(\mathbf{y}) = \sum_{j=1}^{N}\sum_{q=1}^{Q_j}s_{jq, l} \times \eta_{d}\left(\|\mathbf{y} - \mathbf{y}_{jq}\|\right) + \sum_{p = 1}^{d + 1}\alpha_{p, l} \times \phi_{p}(\mathbf{y}), \quad l = 1, 2, \dots, D, \label{eq:3}
\end{equation}
with $\phi_p(\mathbf{y})$ representing the basis of the linear space of polynomials on $\mathbb{R}^{d}$ with degree $\leq 1$ and $\eta_{d}(y) = y^{4-d}\log y$ when $\log y \neq 0$ and $d$ is even, $\eta_{d}(y) = 0$ when $\log y = 0$ and $d$ is even, and $\eta_{d}(y) = y^{4-d}$ when $d$ is odd.

Without accounting for the additive structure of the model, and after concatenating all relevant values into $Q = \sum_{j=1}^{N}Q_{j}$-dimensional vectors of input values $\mathbf{y}$, weights $\mathbf{w}$, and outputs $\mathbf{x}$, we can estimate $f$ by obtaining the values of $s_l \in \mathbb{R}^{Q}$ and $\alpha_l \in \mathbb{R}^{d + 1}$ for $l = 1, \dots, D$ that minimize $\|\mathbf{W}^{\frac{1}{2}}\left(\mathbf{x}_{l} - \mathbf{E}s_l - \mathbf{R}\alpha_l\right)\|^2_{\mathbb{R}^{Q}} + \lambda\|\mathbf{E}^{\frac{1}{2}}s_l\|^2_{\mathbb{R}^{Q}}$ under the constraint that $\mathbf{R}^{T}s_l = 0$, where
\begin{itemize}
  \item $\mathbf{E} = (E_{qr})_{1 \leq q, r, \leq Q}$ is defined by $E_{qr} = \eta_{d}\left(\|\mathbf{y}_{q} - \mathbf{y}_{r}\|\right))$.
  \item $\mathbf{R} = \left(R_{mp}\right)_{1 \leq m \leq Q, \ 1 \leq p \leq d + 1}$ is defined by $R_{mp} = \phi_p(\mathbf{y}_{m})$.
  \item $\mathbf{W} = diag(w_{1}, \dots, w_{Q})$.
\end{itemize}

@@

\begin{equation}
   \left(\begin{array}{ccc}
    2\mathbf{EWE} + 2\lambda\mathbf{E} & 2\mathbf{EWR} & \mathbf{R} \\
    2\mathbf{R}^{T}\mathbf{WE} & 2\mathbf{R}^{T}\mathbf{WR} & \mathbf{0} \\
    \mathbf{R}^{T} & \mathbf{0} & \mathbf{0}
  \end{array}\right)\left(
  \begin{array}{c}
    s_l \\
    \alpha_l \\
    \upsilon_l
  \end{array}
  \right) = \left(
  \begin{array}{c}
    2\mathbf{EW}\mathbf{x}_l \\
    2\mathbf{R}^{T}\mathbf{W}\mathbf{x}_l \\
    \mathbf{0}
  \end{array}
  \right). \label{eq:4}
\end{equation}
The optimal value of $\lambda$ can be found using cross-validation, where with images being randomly split into $K$ folds, functions may be evaluated by the mean squared error, written as
\begin{equation}
  MSD(\lambda) = \frac{1}{K}\sum_{k=1}^{K}\frac{1}{N_u}\sum_{j=1}^{N_u}\frac{1}{Q_j}\sum_{q=1}^{Q_j}\|\mathbf{x}_{jq} - f_{\lambda}^{(k)}(\mathbf{y}_{jq})\|^2. \label{eq:5}
\end{equation}
Here, $f_{\lambda}^{(k)}(\mathbf{y})$ denotes the estimated spline function given tuning parameter $\lambda$ and with the data for the $k$th fold excluded.

The coefficients and tuning parameter that minimize $\mathcal{K}_{\lambda}(f)$ suffice for estimating $f_{j}(\mathbf{y})$, in which the coefficients are not time-varying. To reach estimates for the varying-coefficient models, equation \ref{eq:4} is solved using the data available at each time point within the relevant study level. To estimate $f_p(\mathbf{y}, t)$, we use $\left\{\mathbf{x}_t\right\}_{t=1}^T$ and $\left\{\mathbf{y}_t\right\}_{t=1}^T$, yielding fitted function $\left\{\hat{f}_t(\mathbf{y})\right\}_{t=1}^T$. Using the maximum and minimum values of $\mathbf{y}_t$ in each dimension, a new $d$-dimensional grid of $Q' = \text{max}\left(\left\{Q_{j}\right\}_{j = 1}^{N}\right)$ knots, $\left\{\mathbf{y'}_q\right\}_{q=1}^{Q'}$, is generated with the knots being evenly spread throughout the parameter space. These parameters serve as common knots that allow the spline coefficients to be compared across time points. They are then used to compute a new set of corresponding outcome values $\left\{\mathbf{x}'_t\right\}_{t=1}^{T}$. The spline functions are reestimated as $\left\{\hat{f}'_t\right\}_{t=1}^{T}$ using $\left\{\mathbf{x}'_t\right\}_{t=1}^{T}$ and $\mathbf{y}'$. These newly generated knots are weighted equally, so an unweighted smoothing spline model is used. Following \cite{greenSilverman1994}, we let

\begin{itemize}
  \item $\mathbf{E}' = \left(E'_{qr}\right)_{1 \leq q, r, \leq Q'}$ defined by $E'_{qr} = \eta_{d}\left(\|\mathbf{y}'_{q} - \mathbf{y}'_{r}\|\right)$.
  \item $\mathbf{R}' = \left(R'_{mp}\right)_{1 \leq m \leq Q', \ 1 \leq p \leq d + 1}$ is defined by $R'_{mp} = \phi_p(\mathbf{y'}_{m})$.
\end{itemize}

@@

As with the tuning process for $\lambda$, the optimal value of $\gamma$ can be identified using cross-validation, with either leave-one-out or $k$-folds cross-validation being appropriate depending on the number of time points available. If using leave-one-out cross-validation, the mean squared error can be calculated as
\begin{equation}
  MSD(\gamma) = \frac{1}{T}\sum_{t=1}^{T}\frac{1}{N_t}\sum_{j=1}^{N_t}\frac{1}{Q_j}\sum_{q=1}^{Q_j}\|\mathbf{x}_{jq} - f_{\gamma}^{(t)}(\mathbf{y}_{jq}, t_j)\|^2. \label{eq:8}
\end{equation}
We then let $\hat{\gamma}$ be the value of $\gamma$ that minimizes $MSD(\gamma)$.

With a proper estimate of the $\hat{\gamma}$, we may denote the coefficients of $g(t)$ as $\Omega = \left(\delta^{T}, \tau^{T}\right)$, and those estimated by $g(t)$ as $\omega(t) = \left(\mathbf{s}_{\hat{\gamma}}(t)^{T}, \mathbf{\alpha}_{\hat{\gamma}}(t)^{T}\right)^{T}$. We use these coefficients in equation \ref{eq:3} in place of the constant coefficients. Therefore, the varying-coefficient models have the following outputs:
\begin{equation}
  f(\mathbf{y}, t) = \sum_{q=1}^{Q'}\mathbf{s}_{\hat{\gamma}, l}(t) \eta_{d}\left(\|\mathbf{y} - \mathbf{y'}_m\|\right) + \sum_{p=1}^{d + 1}\mathbf{\alpha}_{\hat{\gamma}, p, l}(t) \phi_p(\mathbf{y}), \quad l = 1, \dots, D. \label{eq:9}
\end{equation}
The fitting process for the varying-coefficient smoothing spline model is given formally as Algorithm \ref{alg:varying_spline}.

@@

\begin{algorithm}
\caption{Varying-Coefficient Smoothing Spline Model}\label{alg:varying_spline}
  \KwData{$D$-dimensional output values $\left\{\mathbf{x}_{jq}\right\}_{j = 1, q = 1}^{N, Q_{j}}$, $d$-dimensional input values $\left\{\mathbf{y}_{jq}\right\}_{j=1, q=1}^{N, Q_{j}}$, weights $\left\{w_{jq}\right\}_{j=1, q=1}^{N, Q_{j}}$, observation times $\left\{t_{j}\right\}_{j=1}^{N}$, candidate tuning parameters $\left\{\lambda_l\right\}_{l=1}^{L}$, $\left\{\gamma_p\right\}_{p=1}^{P}$, and number of cross-validation folds $K$.}
  \KwResult{Analytic formula of $\hat{f}(\mathbf{y}, t)$, coefficients $\Omega = \left(\mathbf{\delta}^{T}, \mathbf{\tau}^{T}\right)$, and optimal tuning parameter $\hat{\gamma}$.}

\For{$t = 1, 2, \dots, T$} {
  Set $U_t$ equal to the number of images included in the data at time $t$\;
  \For{$l = 1, 2, \dots, L$} {
    Solve equation \ref{eq:4} to estimate $f_{\lambda_l, t}(\mathbf{y})$ with inputs $\mathbf{x}_t$, $\mathbf{y}_t$, and $w_t$\;
    \For{$u = 1, 2, \dots, U_t$} {
      Solve equation \ref{eq:4} to estimate $f^{(-u)}_{\lambda_l, t}(\mathbf{y})$ with inputs $\mathbf{x}^{(-u)}_t$, $\mathbf{y}^{(-u)}_t$, $w^{(-u)}_t$\;
      Estimate $MSD(\lambda_{l})_u$ using equation \ref{eq:5}\;
    }
    Set $MSD(\lambda_l) = \frac{1}{U}\sum_{u=1}^{U}MSD(\lambda_l)_u$\;
  }
  $\hat{\lambda}_t = \arg \min_{\lambda} MSD(\lambda)$\;
  $f_t(\mathbf{y}) = f_{\hat{\lambda}_t}(\mathbf{y})$\;
}

  Let $Q' = \max(Q_{j})$, $\left\{\mathbf{y}'_{q}\right\}_{q=1}^{Q'}$ be a grid spanning the range of values of $\left\{\mathbf{y}_{jq}\right\}_{j=1, q=1}^{N, Q_j}$\;
  \For{$t = 1, 2, \dots, T$} {
    Compute $\left\{\mathbf{x'}_{q, t}\right\}_{q=1}^{Q'} = \left\{f_t(\mathbf{y'}_q)\right\}_{q=1}^{Q'}$\;
    Solve equation \ref{eq:6} to compute $\mathbf{s'}_t$ and $\mathbf{\alpha'}_t$\;
    Set $\mathbf{\omega}_t = \left[\mathbf{s'}_t^{T}, \mathbf{\alpha'}_t^{T}\right]$\;
  }
  \For{$p = 1, 2, \dots, P$} {
    Compute $g_{\gamma_p}(t)$ by solving equation \ref{eq:7}\;
    \For{$t = 1, 2, \dots, T$} {
      Compute $g_{\gamma_p}^{(t)}$ by solving \ref{eq:7}\;
    }
    Estimate $MSD(\gamma_p)$ using \ref{eq:8}\;
  }
  $\hat{\gamma} = \arg \min_{\gamma} MSD(\gamma)$\;
  $\hat{f}(\mathbf{y}, t) = f_{\hat{\gamma}}(\mathbf{y}, t)$, where the form of $\hat{f}(\mathbf{y}, t)$ is given in \ref{eq:9}.
\end{algorithm}

@@

\[%
  f_p(\mathbf{y}_{jq}, t_{j}) = E\left[\mathbf{x}_{jq} - f_{g(j)}(\mathbf{y}_{jq}, t_{j}) - f_{i(j)}(\mathbf{y}_{jq}, t_{j}) - f_{j}(\mathbf{y}_{jq})\right]
,\]%
with a similar result holding for each of the smooth functions making up the additive model. The model in its entirety is considered to have converged when the maximum of the distances between the coefficients of functions estimated for the same element in consecutive steps is below a given threshold. Details of the backfitting algorithm to fit the proposed model is given in Algorithm \ref{alg:backfitting}.

\begin{algorithm}
\footnotesize
\caption{Backfitting Algorithm for Additive Spline Model}\label{alg:backfitting}
  \KwData{$D$-dimensional output values $\left\{\mathbf{x}_{jq}\right\}_{j=1, q=1}^{N, Q_{j}}$, $d$-dimensional input values $\left\{\mathbf{y}_{jq}\right\}_{j=1, q=1}^{N, Q_{j}}$, observation times $\left\{t_{j}\right\}_{j=1}^{N}$, weights $\left\{w_{jq}\right\}_{j=1, q=1}^{N, Q_j}$, group and individual indicator values $\left\{g(j)\right\}_{j=1}^{N}$ and $\left\{i(j)\right\}_{j=1}^{N}$, candidate tuning parameters $\left\{\lambda_l\right\}_{l=1}^{L}$ and $\left\{\gamma_p\right\}_{p=1}^{P}$, $\epsilon$ and $K$.}
  \KwResult{Analytic formulas of $\hat{f}_p$, $\left\{\hat{f}_g\right\}_{g=1}^{G}, \left\{\hat{f}_i\right\}_{i=1}^{I}$, and $\left\{\hat{f}_{j}\right\}_{j=1}^{N}$.}

  Estimate $f_{p}^{(0)}(\mathbf{y}, t)$ and $\Omega_{p}^{(0)}$ using Algorithm \ref{alg:varying_spline} with inputs $\left\{\mathbf{x}_{jq}\right\}_{j=1, q=1}^{N, Q_j}$, $\left\{\mathbf{y}_{jq}\right\}_{j=1, q=1}^{N, Q_j}$, $\left\{w_{jq}\right\}_{j=1, q=1}^{N, Q_j}$, $\left\{t_{j}\right\}_{j=1}^{N}$, $\left\{\lambda_l\right\}_{l=1}^{L}$, $\left\{\gamma_p\right\}_{p=1}^{P}$, and $K$\;
